Typed Functional Genetic Programming

### 2013-06-24 tex/thesis.tex
[tex] GP outline

\subsection{Generating initial population}

@@

\textbf{TODO!}

@@

For John Koza the most important genetic operator in GP is 
crossover. It is operator inspired by sexual reproduction
occurring in the nature. Generally speaking crossover takes
two (parent) individuals and combines theirs genomes to produce 
two possibly new (child) individuals.

@@

\subsection{Reproduction}
\subsection{Mutation}

@@

Here is pseudocode for GP algorithm:

\Pseudokod{GP(fitness,buildingBlocks,numGens,popSize,probabs)}{
	$gen \leftarrow$ 0 \;
	$pop \leftarrow generateInitialPopuletion( buildingBlocks ) $ \;
	($popWithF,terminate,best) \leftarrow$ evaluate($fitness$, $pop$)\;

@@

	\While{ $i < popSize$ }{
		$op \leftarrow probabilisticallySelectOperation(probabs)$ \;		
		\Switch{op}{
			\Case{Crossover}{
				$parent1 \leftarrow$ selection( $popWithF$ ) \;
				$parent2 \leftarrow$ selection( $popWithF$ ) \;
				($child1$,$child2$) = crossover( $parent1$ , $parent2$ ) \;
				$newPop$.insert( $child1$ ) \;
				$newPop$.insert( $child2$ ) \;
				$i \leftarrow i + 2$ \;
			}
			\Case{Reproduction}{
				$indiv \leftarrow$ selection( $popWithF$ ) \;
				$newPop$.insert( $indiv$ ) \;
				$i \leftarrow i + 1$ \;
			}
			\Case{Mutation}{
				$indiv \leftarrow$ selection( $popWithF$ ) \;
				$mutant \leftarrow$ mutate( $indiv$ ) \;
				$newPop$.insert( $mutant$ ) \;
				$i \leftarrow i + 1$ \;			
			}		
		}
	}

	$pop \leftarrow newPop$  \;
	($popWithF,terminate,best) \leftarrow$ evaluate($fitness$, $pop$)\;
	$gen \leftarrow gen + 1$ \; 
 }
 \Return pop \;
}

@@

@@

\section{Lambda term}
\label{deflam}

>>>>

\subsection{Generating individual}

@@

\textbf{TODO!}\\
\textbf{Talk about Generating initial population}

@@

For Koza the most important genetic operator in GP is 
crossover. It is operator inspired by sexual reproduction
occurring in the nature. Generally speaking crossover takes
two (parent) individuals and combines theirs genomes to produce 
two possibly new (child) individuals.

@@

\subsection{Reproduction}

Reproduction is simple mechanism providing preservation of solutions
from the current population to the next one. It simply copies 
one individual to the next population.

\subsection{Mutation}

Mutation is genetic operator modifying one individual.
There are many options for mutation mechanisms 
for us to choose from. Here will be described 
\textit{Subtree generating mutation} 
witch uses mechanism for generating individuals.

This mutation randomly selects one node in the individual tree.
New mutant individual is constructed by replacement of 
subtree with root in the selected node by new generated
tree. This new tree is generated by mechanism for generating 
individuals.

\subsection{Construction of a next population}

Let $pop_{t}$ be the current population which we want to 
transform into the next population $pop_{t+1}$. 
We start by initializing $pop_{t+1}$ by empty population.

Then we iteratively fill $pop_{t+1}$ by individuals 
returned by genetic operators.

In each iteration one genetic operator is randomly selected.
Then required amount of individuals for the operation is selected
by individual selection mechanism described above (one or two
individuals for our genetic operators).
And those selected individuals are used as input for selected
genetic operator which produces some new individuals.
Those new individuals are inserted into $pop_{t+1}$.

This process continues until $pop_{t+1}$ is filled with
$popSize$ individuals.

Selection of genetic operation in each operation is 
controlled by probabilities for each genetic operator.

Koza in \cite{koza92} uses as default probabilities those values:

\begin{itemize}
	\item \textit{Crossover}    : $70\%$
	\item \textit{Reproduction} : $30\%$
	\item \textit{Mutation}     :  $0\%$
\end{itemize}

@@

Bellow is described GP algorithm in pseudocode.
Inputs are:

\begin{itemize}
	\item \textit{fitness} - Fitness function.
	\item \textit{TuF} - Building blocks accompanied with an  	
	      information about implementations and argument numbers.
	\item \textit{popSize} - Population size.
	\item \textit{numGens} - Number of generations.
	\item \textit{probabs} - Genetic operators probabilities.
\end{itemize}

\Pseudokod{GP(fitness,TuF,popSize,numGens,probabs)}{
	$gen \leftarrow$ 0 \;
	$pop \leftarrow generateInitialPopuletion( TuF ) $ \;
	($popWithF,terminate,best) \leftarrow$ 
	evaluate($fitness$, $TuF$, $pop$)\;

@@

	\While{ $i < popSize$ }{
		$op \leftarrow probabilisticallySelectOperation(probabs)$ \;		
		\Switch{op}{
			\Case{Crossover}{
				$parent1 \leftarrow$ selection( $popWithF$ ) \;
				$parent2 \leftarrow$ selection( $popWithF$ ) \;
				($child1$,$child2$) = crossover( $parent1$ , $parent2$ ) \;
				$newPop$.insert( $child1$ ) \;
				$newPop$.insert( $child2$ ) \;
				$i \leftarrow i + 2$ \;
			}
			\Case{Reproduction}{
				$indiv \leftarrow$ selection( $popWithF$ ) \;
				$newPop$.insert( $indiv$ ) \;
				$i \leftarrow i + 1$ \;
			}
			\Case{Mutation}{
				$indiv \leftarrow$ selection( $popWithF$ ) \;
				$mutant \leftarrow$ mutate( $indiv$ , $TuF$ ) \;
				$newPop$.insert( $mutant$ ) \;
				$i \leftarrow i + 1$ \;			
			}		
		}
	}

	$pop \leftarrow newPop$  \;
	($popWithF,terminate,best) \leftarrow$ 
	evaluate($fitness$, $TuF$, $pop$)\;
	$gen \leftarrow gen + 1$ \; 
 }
 \Return pop \;
}

@@

\textbf{TODO mention best preservation and same popSize checking. }

@@

\newpage		
\section{Lambda term}
\label{deflam}
